---
title: "Comparing different skewness calculations"
output: github_document
---


Do different methods for calculating skewness give qualitatively similar results?

(For later - evenness?)

```{r setup}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(scads)
library(ggplot2)
dataset <- "large"
```

```{r get an SAD}
if(dataset == "plants") {
  
  portal_sad <- portalr::plant_abundance(level = "Treatment", type = "Winter Annuals", plots = "All", unknowns = F, correct_sp = T, shape = "flat", min_quads = 16) %>%
    filter(treatment == "control") %>%
    select(year, species, abundance) %>%
    filter(year == 1994) %>%
    select(-year) %>%
    rename(abund = abundance) %>%
    select(abund) %>%
    arrange(abund) %>%
    as.matrix() %>%
    as.vector()
  

} else if(dataset == "rodents94") {
  portal_data <- isds::get_toy_portal_data(download=T, years = c(1994))
  
  portal_sad <- portal_data %>%
    select(species) %>%
    group_by(species) %>%
    summarize(abund = n()) %>%
    ungroup() %>%
    select(abund) %>%
    arrange(abund) %>%
    as.matrix() %>%
    as.vector()
  

} else {
  portal_data <- isds::get_toy_portal_data(download=T, years = c(1990:1995))
  
  portal_sad <- portal_data %>%
    select(species) %>%
    group_by(species) %>%
    summarize(abund = n()) %>%
    ungroup() %>%
    select(abund) %>%
    arrange(abund) %>%
    as.matrix() %>%
    as.vector()
  
}

nspp = length(portal_sad)
nind = sum(portal_sad)

portal_sad <- data.frame(
  abund = portal_sad,
  source = "emp",
  rank = 1:length(portal_sad),
  stringsAsFactors = F
)

```

```{r get a BUNCH of fs samples, eval = F, include=T}

fs_bank <- sample_feasibleset(s = nspp, n = nind, nsamples = 10000, distinct = TRUE)
save(fs_bank, file = "fs_bank_94.Rds")

```

```{r load fs bank}

if(dataset == "plants") {
  load(here::here("fs_bank_plants.Rds"))
  fs_bank <- fs_bank[, 1:10000]

  nametoprint <- "Portal winter annuals 1994, 10000 draws"
} else if(dataset == "rodents94") {
  load(here::here("fs_bank_94.Rds"))
  nametoprint <- "Portal control rodents 1994, 10000 draws"
} else if(dataset == "small") {
  load(here::here("fs_bank_small.Rds"))
  nametoprint <- "Portal control rodents 1990-95, 100 draws"
} else {
  load(here::here("fs_bank.Rds"))
  nametoprint <- "Portal control rodents 1990-95, 10000 draws"
}

```

```{r manip fs_bank}

fs_bank_M <- fs_bank %>%
  as.data.frame(stringsAsFactors = F) %>%
  mutate(rank = row_number()) %>%
  tidyr::gather(-rank, key = "sim", value = "abund") %>%
  mutate(sim = as.integer(substr(sim, 2, nchar(sim))))

```


# This report is for `r nametoprint`. 

## Density plots of raw rank abundances

The y-axis is abundance. Each black dot is an abundance value from a vector drawn from the feasible set. The red line plots the distribution from Portal.

```{r plot rads and rescaled rads, fig.width = 10}
fs_dp <- ggplot(data = fs_bank_M, aes(x =as.factor(rank), y = abund)) + 
  geom_point(alpha = .2, size = 1) +
  geom_line(data = portal_sad, aes(x = rank, y = abund), alpha = .5, color = "red") +
  theme_bw() +
  ggtitle("Abundance: draws from FS vs Portal")

fs_dp
```

## Skewness algorithms

`e1071` has three algorithms available for skewness. Do they give the same results?


```{r skewness algs, fig.width = 12}
fs_bank_M <- fs_bank_M %>%
  mutate(source = "sampled") %>%
  bind_rows(mutate(portal_sad, sim = -99))


fs_bank_skews <- fs_bank_M %>%
  group_by(sim, source) %>%
  summarize(skew_1 = e1071::skewness(abund, type = 1),
            skew_2 = e1071::skewness(abund, type = 2),
            skew_3 = e1071::skewness(abund, type = 3)) %>%
  ungroup()# %>%
  #tidyr::gather(-sim, -source, key = "skew_alg", value = "skew_val")


skew_plots <- list()

skew_plots[[1]] <- ggplot(data = fs_bank_skews, aes(x = skew_1, y = skew_2)) +
  geom_point(alpha = .05) +
  theme_bw() +
  ggtitle("Skew 1 v. Skew 2")

skew_plots[[2]] <- ggplot(data = fs_bank_skews, aes(x = skew_1, y = skew_3)) +
  geom_point(alpha = .05) +
  theme_bw() +
  ggtitle("Skew 1 v. Skew 3") 

skew_plots[[3]] <- ggplot(data = fs_bank_skews, aes(x = skew_2, y = skew_3)) +
  geom_point(alpha = .05) +
  theme_bw() +
  ggtitle("Skew 2 v. Skew 3")

gridExtra::grid.arrange(grobs = skew_plots, nrow = 1)

```

```{r ranked skew, fig.width = 12}

fs_bank_skews <- fs_bank_skews %>%
  arrange((skew_1)) %>%
  mutate(skew_1_rank = row_number()) %>%
  arrange((skew_2)) %>%
  mutate(skew_2_rank = row_number()) %>%
  arrange((skew_3)) %>%
  mutate(skew_3_rank = row_number())

fs_bank_rank_plots <- list()

fs_bank_rank_plots[[1]] <- ggplot(data = fs_bank_skews, aes(x = skew_1_rank, y = skew_2_rank)) +
  geom_point(alpha = .05) +
  theme_bw() +
  ggtitle("Skew 1 v. Skew 2 Rank")

fs_bank_rank_plots[[2]] <- ggplot(data = fs_bank_skews, aes(x = skew_1_rank, y = skew_3_rank)) +
  geom_point(alpha = .05) +
  theme_bw() +
  ggtitle("Skew 1 v. Skew 3 Rank") 

fs_bank_rank_plots[[3]] <- ggplot(data = fs_bank_skews, aes(x = skew_2_rank, y = skew_3_rank)) +
  geom_point(alpha = .05) +
  theme_bw() +
  ggtitle("Skew 2 v. Skew 3 Rank")


gridExtra::grid.arrange(grobs = fs_bank_rank_plots, nrow = 1)
```

Are literally all of the ranks exactly the same?

```{r literally equal, include = T}
all((fs_bank_skews$skew_1_rank == fs_bank_skews$skew_2_rank),
    (fs_bank_skews$skew_1_rank == fs_bank_skews$skew_3_rank),
    (fs_bank_skews$skew_3_rank == fs_bank_skews$skew_2_rank))

```



## Diversity indices

Moving forward, we'll just use one skewness alogrithm. Comparing the ranks for skewness and diversity indices calculated using `vegan`:

* Skewness
* Shannon-Weiner diversity, natural log base
* Simpson's index (1-D)
* Fisher's alpha (not bringing this one forward - for fixed S and N, alpha does not change. This is expected because the logseries is entirely determined by S and N)


```{r add div indices}
library(vegan)

fs_bank_div <- fs_bank_M %>%
  group_by(sim, source) %>%
  summarize(
    skew = e1071::skewness(abund),
    shannon = vegan::diversity(abund, index = "shannon", base = exp(1)),
    simpson = vegan::diversity(abund, index = "simpson"),
    falpha = vegan::fisher.alpha(abund)
  ) %>%
  ungroup() %>%
  mutate_at(c("skew", "shannon", "simpson", "falpha"), list(rank = dense_rank)) 

fs_bank_div_long <- fs_bank_div %>%
  tidyr::gather(-sim, -source, key = "variable", value = "val")
```


### Diversity index 1:1 plots
```{r rank 1 to 1 plots, fig.height = 4, fig.width = 12}

skew_shannon <- ggplot(data = fs_bank_div, aes(x = skew_rank, y = shannon_rank)) +
  geom_point(alpha = .05) +
  geom_abline(intercept = max(fs_bank_M$sim), slope = -1) +
  theme_bw() +
  ggtitle("Skewness vs. Shannon")

skew_simpson <-  ggplot(data = fs_bank_div, aes(x = skew_rank, y = simpson_rank)) +
  geom_point(alpha = .05) +
  geom_abline(intercept =  max(fs_bank_M$sim), slope = -1) +
  theme_bw() +
  ggtitle("Skewness vs. Simpson")

shannon_simpson <-  ggplot(data = fs_bank_div, aes(x = shannon_rank, y = simpson_rank)) +
  geom_point(alpha = .05) +
  geom_abline(intercept = 0, slope = 1) +
  theme_bw() +
  ggtitle("Shannon vs. Simpson")

onetoones <- gridExtra::grid.arrange(grobs = list(skew_shannon, skew_simpson, shannon_simpson), ncol = 3)

onetoones

```

There is a general, but not perfect, relationship between skewness and other diversity indices. 


### Position of empirical within sim distributions

```{r sim emp plots, fig.width = 4, fig.height = 12, fig.align="center"}

sim_emp_plots <- ggplot(data = filter(fs_bank_div_long, variable %in% c("skew", "shannon", "simpson")), aes(x = 0, y = val, color = source)) +
  geom_violin(data = filter(fs_bank_div_long,  variable %in% c("skew", "shannon", "simpson"), source == "sampled")) +
  geom_point(data = filter(fs_bank_div_long,  variable %in% c("skew", "shannon", "simpson"), source == "emp"), size = 2) +
  theme_bw() +
  scale_color_viridis_d(end = .8) +
  facet_wrap(vars(variable), ncol = 1, scales = "free", strip.position = "top")

sim_emp_plots
```

The position of the empirical is generally qualitatively consistent but there is variation. I think the next step is to add diversity indices to the analysis at scale in diazrenata/scadsplants. It is not very computationally expensive to bring 2 additional indices through the full analytical pipeline, and including traditional metrics alongside skewness makes this more intuitive to an ecological background. 